\vspace{-3em} \textbf{ Mapping stellar surfaces\\ II: An interpretable Gaussian process model for light curves }

### 2021-01-28 paper2/ms.tex

However, understanding the properties of stellar surfaces and starspots
in particular is a crucial step
toward understanding stellar magnetism, which plays a fundamental part
in stellar interior structure and evolution. Stellar magnetic fields
control the spin-down of stars over time, on which the field of
gyrochronology is founded \citep[][]{Barnes2001,Angus2019}. They affect
wave propagation in stellar interiors and must be properly
understood to interpret asteroseismic measurements \citep[e.g.,][]{Fuller2015}.
Strong magnetic fields are also likely the driving force behind chemical
peculiarity in Ap/Bp stars~\citep{Turcotte2003,Sikora2018}, as well as
radius inflation in M dwarfs \citep{Gough1966,Ireland2018}.
Significant effort has thus been spent on theoretical and numerical approaches to understanding
magnetohydrodynamics (MHD) in stars~\citep[e.g.,][]{Miesch2009}.
Theoretical progress is challenging; MHD equations are nonlinear, since
magnetism often emerges in regions of highly developed turbulence. There is a
large separation of scales between the size of a star and the
magnetic/thermal/momentum diffusion length-scales, and so numerical
simulations at realistic resolutions remain out of reach. As a result
there is still significant theoretical uncertainty over how dynamos
manifest in stars of different masses, how magnetism interacts with rotation,
and how star spots form \citep{Yadav2015,Weber2016}. Because of the
significant impacts of magnetism on every branch of stellar astrophysics,
these difficulties make new constraints from observations especially valuable.

>>>>

\xxx{WORK IN PROGRESS:}

However, understanding the properties of stellar surfaces and starspots
in particular is a crucial step
toward understanding stellar magnetism, which plays a fundamental part
in stellar interior structure and evolution. Stellar magnetic fields
control the spin-down of stars over time, on which the field of
gyrochronology is founded \citep[][]{Barnes2001,Angus2019}. They affect
wave propagation in stellar interiors and must be properly
understood to interpret asteroseismic measurements \citep[e.g.,][]{Fuller2015}.
Strong magnetic fields are also likely the driving force behind chemical
peculiarity in Ap/Bp stars~\citep{Turcotte2003,Sikora2018}, as well as
radius inflation in M dwarfs \citep{Gough1966,Ireland2018}.
Stellar magnetohydrodynamics (MHD) is therefore an active area of
research \citep[e.g.,][]{Miesch2009,XXX}.
%
But because of the nonlinearity of the MHD equations and the vast range
of scales on which magnetic processes operate, theoretical progress
is difficult; there is still significant theoretical uncertainty over how dynamos
manifest in stars of different masses, how magnetism interacts with rotation,
and how star spots form \citep{Yadav2015,Weber2016}. Because of the
significant impacts of magnetism on every branch of stellar astrophysics,
these difficulties make new constraints from observations especially valuable.
